---
title: "Assignment1: Reproducible Research"
output: html_document
---
#Package use
```{r Include packages}
library(knitr)
library(dplyr)
library(ggplot2)
```
#Loading and Preprocessing the Data
```{r Include data1}
data_full <- read.csv("C:/Users/BTM KPKT/Documents/GitHub/Reproducible Research/Assignment1 Reproducible/activity.csv", header=T,na.strings="NA", nrows=17568)
data_clean <- data_full[with(data_full, {!(is.na(steps))}),]
head(data_clean)
```
#What is mean total number of steps taken per day?
##Calculate the total number of steps taken per day
```{r Include data2}
by_date <- group_by(data_clean, date)
steps_by_date <- summarise(by_date, total = sum(steps))
steps_by_date
```

##Make a histogram of the total number of steps taken each day
```{r Include histogram}
hist(steps_by_date$total, main="Histogram of total number of steps per day", 
     xlab="Total number of steps in a day")
```

##Calculate and report the mean and median of the total number of steps taken per day
```{r Include data3}
summary(steps_by_date)
###Mean = 10766; Median = 10765
```

#What is the average daily activity pattern?
```{r Include data4}
steps_by_interval <- aggregate(steps ~ interval, data_clean, mean)
head(steps_by_interval)
```


```{R includes scatterplot}
plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
     main="Average number of steps over all days", xlab="Interval", 
     ylab="Average number of steps")
```

##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{R Include data5}
# find row with max of steps
max_steps_row <- which.max(steps_by_interval$steps)

# find interval with this max
steps_by_interval[max_steps_row, ]

## Interval 835 contains the maximum number of steps
```

#Imputing missing values
##Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{R Include data6}
sum(is.na(data_full))
put_a_data <- data_full
for (i in 1:nrow(put_a_data)) {
        if (is.na(put_a_data$steps[i])) {
                interval_value <- put_a_data$interval[i]
                steps_value <- steps_by_interval[
                        steps_by_interval$interval == interval_value,]
                put_a_data$steps[i] <- steps_value$steps
        }
}
###NA is replaced by mean of 5 minute-interval
head(put_a_data)
###New data set with no NA.
```

```{R Include data7}
by_date_new <- group_by(put_a_data, date)
steps_by_date2 <- summarise(by_date_new, total = sum(steps))
steps_by_date2
```

##Make a histogram of the total number of steps taken each day
```{R Include histogram2}
hist(steps_by_date2$total, main="Histogram of total number of steps per day (no missing value)", 
     xlab="Total number of steps in a day")
```
##Calculate and report the mean and median total number of steps taken per day. 
# get mean and median of new data set (with no missing value)
```{R Include the data8}
summary(steps_by_date)
summary(steps_by_date2)
###Mean is the same.  Median is different.
###The impact of imputing missing data on the estimates of the total daily number of steps will provide different result.  
###Thus we must choose the correct method to impute the data.
```

#Are there differences in activity patterns between weekdays and weekends?
##Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{R Include new data9}
put_a_data['type_of_day'] <- weekdays(as.Date(put_a_data$date))
head(put_a_data)
put_a_data$type_of_day[put_a_data$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
put_a_data$type_of_day[put_a_data$type_of_day != "weekend"] <- "weekday"
head(put_a_data)
```
##Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken.
```{R Include data10}
# convert type_of_day from character to factor
put_a_data$type_of_day <- as.factor(put_a_data$type_of_day)
str(put_a_data)
# calculate average steps by interval across all days
put_a_data_steps_by_interval <- aggregate(steps ~ interval + type_of_day, put_a_data, mean)
```

# create a plot
```{R Include plot2}
qplot(interval, 
      steps, 
      data = put_a_data_steps_by_interval, 
      type = 'l', 
      geom=c("line"),
      xlab = "Interval", 
      ylab = "Number of steps", 
      main = "") +
        facet_wrap(~ type_of_day, ncol = 1)
```